Fast, Scalable Phrase-Based SMT Decoding

### 2016-05-04 acl.2016/acl2016.tex
changes

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding time with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.

We will focus on two techniques. One, the default translation rule caching mechanism in Moses saves the most recently used rules. However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}

@@

Similar to the phrase-table, the lexicalized reordering model is trained on parallel data where a model file is created which is then queried during decoding. Inevitably, the querying results in loss of decoding speed. Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase querying speed.

However, the model's query key are the source and target phrase of each translation rule. Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

@@

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

>>>>

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding time with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.

We propose two optimizations. One, the default translation rule caching mechanism in Moses saves the most recently used rules. However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}

@@

Similar to the phrase-table, the lexicalized reordering model is trained on parallel data. A resultant model file is created in training which is then queried during decoding. Inevitably, the querying results in loss of decoding speed. Previous work such as~\newcite{junczys_tsd_2012b} improve querying speed with more efficient data structures.

However, the model's query keys are the source and target phrase of each translation rule. Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

@@

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact data structure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}
